---
title: Testing Multivariate simulation for decreasing response variance
author: Raju Rimal
runtime: shiny
output: 
  html_document:
    css: custom.css
    theme: cosmo
---



# Simulation Parameters
```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = NULL,
  echo = FALSE
)
library(shiny)
library(simrel)
```

```{r}
fluidRow(
  column(3, numericInput("n", "Number of observation", 10000, step = 50, min = 10, width = "100%")),
  column(3, numericInput("p", "Number of predictors", 10, min = 2, width = "100%")),
  column(3, numericInput("m", "Number of responses", 3, min = 2, width = "100%")),
  column(2, numericInput("seed", "seed", 123, min = 2, width = "100%")),
  column(1, actionButton('simulate', label = NULL, class="btn-primary", icon = icon("refresh"), width = '100%')),
  column(4, textInput("q", "Number of relevant predictors", value = "5, 4", width = "100%")),
  column(4, textInput("relpos", "Position of Relevant components", value = "1, 2; 3, 4, 6", width = '100%')),
  column(4, textInput("ypos", "Position of response components", value = "1; 2, 3", width = '100%')),
  column(4, sliderInput("gamma", "Decay factor of eigenvector of predictors", 
                        value = 0.5, min = 0, max = 3, width = '100%', step = 0.01)),
  column(4, sliderInput("eta", "Decay factor of eigenvector of response", 
                        value = 0.5, min = 0, max = 3, width = '100%', step = 0.01)),
  column(4, textInput("R2", "Coefficient of Determination", value = "0.7, 0.8", width = '100%'))
)
evl <- function(x) {
  out <- lapply(strsplit(unlist(strsplit(x, ";")), ","), as.numeric)
  if (length(out) == 1) out <- out[[1]]
  return(out)
}
opts <- reactive({
  list(
    n = input$n,
    p = input$p,
    q = evl(input$q),
    relpos = evl(input$relpos),
    gamma = input$gamma,
    R2 = evl(input$R2),
    ypos = evl(input$ypos),
    m = input$m,
    eta = input$eta,
    type = "multivariate"
  )
})
```

## Simulation and Modeling {.tabset}

```{r}
sobj <- eventReactive(input$simulate, {
  set.seed(input$seed)
  do.call(simrel, opts())
})
mdl <- reactive({
  dta <- data.frame(
    y = I(sobj()[["Y"]]),
    x = I(sobj()[["X"]])
  )
  return(lm(y ~ x, data = dta))
})
mdl1 <- reactive({
  dta <- data.frame(
    w <- I(sobj()[["W"]]),
    z <- I(sobj()[["Z"]])
  )
  return(lm(w ~ z, data = dta))
})
```

### Covariance of Responses {.colset}

#### Covariance of simulated Response {.cols}

```{r}
column(
  width = 12, 
  renderPrint({
    round(var(sobj()$Y), 4)
  })
)
```

#### True covariance of response {.cols}

```{r}
column(
  width = 12, 
  renderPrint({
  sigma <- sobj()$Sigma
  m <- sobj()$m
  round(sigma[1:m, 1:m], 4)
})
)
```

### Coefficient of Determination of W {.colset}

#### Estimated Coefficient of determination {.cols}

$$\widehat{\rho_w^2} = \text{sign}(\text{cor}(\mathbf{w}, \mathbf{\widehat{w}})) \times \left[\text{cor}(\mathbf{w}, \mathbf{\widehat{w}})\right]^2$$

```{r}
column(
  width = 12,
    renderPrint({
    fittedW <- fitted(mdl1())
    trueW <- sobj()[["W"]]
    cormat <- cor(trueW, fittedW)
    sgn <- sign(cormat)
    round(cormat^2 * sgn, 4)
  })
)
```

#### Population Coefficient of determination {.cols}

$$\left(\rho_w^2\right)_{ij} = \left(
\frac{\sigma_{zw_i}^t\Sigma_{zz}^{-1}\sigma_{zw_i}}{\sigma_{w_{ii}}\sigma_{w_{jj}}}
\right) \\\forall i \text{ (row) and }j\text{ (column) } = 1, 2, \ldots, m$$

```{r}
column(
  width = 12,
    renderPrint({
    sobj()[["RsqW"]]
  })
)
```


#### R-squared values obtained from linear model W ~ Z {.cols}


```{r}
column(
  width = 12,
  renderPrint({
    sapply(summary(mdl1()), function(x) x[["r.squared"]])
  })
)
```

#### Explained to Total Ratio {.cols}
$$(\widehat{\rho}_w)_{ij} = \frac{\text{cov}(\widehat{w}_i \widehat{w}_j)}{\sqrt{\text{var}(w_i) \times \text{var}(w_j)}} \\\forall i \text{ (row) and }j\text{ (column) } = 1, 2, \ldots, m$$
```{r}
column(
  width = 12,
  renderPrint({
    W <- sobj()[["W"]]
    fitted_w <- fitted(mdl1())
    cov_fitted_w <- cov(fitted_w)
    cov_w <- cov(W)
    m <- sobj()[["m"]]
    out <- matrix(0, m, m)
    for (row in 1:m){
      for (col in 1:m) {
        out[row, col] <- cov_fitted_w[row, col] / sqrt(cov_w[row, row] * cov_w[col, col])
      }
    }
    out
  })
)
```




### Coefficient of Determination of Y {.colset}

#### Estimated Coefficient of determination {.cols}

$$\widehat{\rho_y^2} = \text{sign}(\text{cor}(\mathbf{y}, \mathbf{\widehat{y}})) \times \left[\text{cor}(\mathbf{y}, \mathbf{\widehat{y}})\right]^2$$

```{r}
column(
  width = 12,
    renderPrint({
    fittedY <- fitted(mdl())
    trueY <- sobj()[["Y"]]
    cormat <- cor(trueY, fittedY)
    sgn <- sign(cormat)
    round(cormat^2 * sgn, 4)
  })
)
```

#### Population Coefficient of determination {.cols}

$$\left(\rho_y^2\right)_{ij} = \left(
\frac{\sigma_{xy_i}^t\Sigma_{xx}^{-1}\sigma_{xy_i}}{\sigma_{y_{ii}}\sigma_{y_{jj}}}
\right) \\\forall i \text{ (row) and }j\text{ (column) } = 1, 2, \ldots, m$$

```{r}
column(
  width = 12,
    renderPrint({
    sobj()[["RsqY"]]
  })
)
```


```{r}
column(
  width = 12,
    renderPrint({
    sobj()[["RsqYalt"]]
  })
)
```


#### R-squared values obtained from linear model Y ~ X {.cols}

```{r}
column(
  width = 12,
  renderPrint({
    sapply(summary(mdl()), function(x) x[["r.squared"]])
  })
)
```


#### Explained to Total Ratio {.cols}
$$(\widehat{\rho}_y)_{ij} = \frac{\text{cov}(\widehat{y}_i \widehat{y}_j)}{\sqrt{\text{var}(y_i) \times \text{var}(y_j)}} \\\forall i \text{ (row) and }j\text{ (column) } = 1, 2, \ldots, m$$
```{r}
column(
  width = 12,
  renderPrint({
    Y <- sobj()[["Y"]]
    fitted_y <- fitted(mdl())
    cov_fitted_y <- cov(fitted_y)
    cov_y <- cov(Y)
    m <- sobj()[["m"]]
    out <- matrix(0, m, m)
    for (row in 1:m){
      for (col in 1:m) {
        out[row, col] <- cov_fitted_y[row, col] / sqrt(cov_y[row, row] * cov_y[col, col])
      }
    }
    out
  })
)
```



